WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2014-08-11 ms.tex
Fix typo

The Fisher information matrix for a measurement $x$ described by the unknown parameter vector $\boldsymbol{\theta}$ is
%
\begin{equation}\label{eq:general-fisher-matrix}
    \mathcal{I}_{jk} = \mathrm{E} \, \left[
        -\frac{\partial^2 \log
            \mathcal{L}(x_i ; \boldsymbol\theta)}
            {\partial \theta_j \theta_k}
    \middle| \boldsymbol\theta
    \right].
\end{equation}
%
In other words, it is the expectation value, condition on the true parameter values, of the Jacobian of the log likelihood. This equation involves double derivatives, but substantial simplification is possible when---as in this case---the likelihood is Gaussian:
%
\begin{equation}\label{eq:gaussian-fisher-matrix}
    \mathcal{I}_{jk} = \int_0^\infty \Re \left[
        \left(\frac{\partial h_i}{\partial \theta_j}\right)^*
        \left(\frac{\partial h_i}{\partial \theta_k}\right)
    \right] \frac{1}{S_i(\omega)} \, d\omega.
\end{equation}
%
This form is more useful because it involves only first derivatives, and of the signal $h (\omega)$ rather than the entire observation $x (\omega)$. In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
%
\begin{equation}\label{eq:angular-frequency-moments}
    {\overline{\omega^k}}_i =
        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \omega^k \, d\omega \right]
        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \, d\omega \right]^{-1},
\end{equation}
%
the Fisher matrix for the signal in the $i$th detector is
%
\begin{equation}
    \mathcal{I}_i = \left(
        \begin{array}{cc}
            \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i} &
            \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_\mathrm{in}} \\
            {\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_\mathrm{in}}}^\intercal &
            {\rho_i}^2 \mathcal{I}_{\boldsymbol\theta_\mathrm{in},\boldsymbol\theta_\mathrm{in}}
        \end{array}
    \right)
\end{equation}
%
where
%
\begin{equation}\label{eq:fisher-matrix}
    \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i} = \bordermatrix{
        ~ & \rho_i & \gamma_i & \tau_i \cr
        \rho_i & 1 & 0 & 0 \cr
        \gamma_i & 0 & {\rho_i}^2 & -{\rho_i}^2 {\overline{\omega}}_i \cr
        \tau_i & 0 & -{\rho_i}^2 {\overline{\omega}}_i & {\rho_i}^2 {\overline{\omega^2}}_i
    }.
\end{equation}
%
(See also \cite{Grover:2013}.) To lowest post\nobreakdashes-Newtonian order, the intrinsic parameters affect the phase evolution but not the amplitude evolution, so for BNS systems we can assume $\mathcal{I}_{\rho_i,\boldsymbol\theta_\mathrm{in}} \approx 0$.
%
\marginpar{Show this?}%
%
When we form the Fisher matrix for the whole network, if the detectors all have the same noise \acp{PSD}, $S_1(\omega) = S_1(\omega) = \cdots = S_n(\omega) \equiv S(\omega)$, then we can change variables from phases and times on arrival $\left\{\gamma_i, \tau_i\right\}_i$ to appropriate differences between them, in order to cancel out all off-diagonal terms between the intrinsic and extrinsic parameters. Because all existing ground-based GW detectors have similar bandpasses, we restrict all of our further attention to just the extrinsic part of the Fisher matrix, $\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i}$.

>>>>

The Fisher information matrix for a measurement $x$ described by the unknown parameter vector $\boldsymbol{\theta}$ is
%
\begin{equation}\label{eq:general-fisher-matrix}
    \mathcal{I}_{jk} = \mathrm{E} \, \left[
        -\frac{\partial^2 \log
            \mathcal{L}(x_i ; \boldsymbol\theta)}
            {\partial \theta_j \theta_k}
    \middle| \boldsymbol\theta
    \right].
\end{equation}
%
In other words, it is the expectation value, condition on the true parameter values, of the Jacobian of the log likelihood. This equation involves double derivatives, but substantial simplification is possible when---as in this case---the likelihood is Gaussian:
%
\begin{equation}\label{eq:gaussian-fisher-matrix}
    \mathcal{I}_{jk} = \int_0^\infty \Re \left[
        \left(\frac{\partial h_i}{\partial \theta_j}\right)^*
        \left(\frac{\partial h_i}{\partial \theta_k}\right)
    \right] \frac{1}{S_i(\omega)} \, d\omega.
\end{equation}
%
This form is more useful because it involves only first derivatives, and of the signal $h (\omega)$ rather than the entire observation $x (\omega)$. In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
%
\begin{equation}\label{eq:angular-frequency-moments}
    {\overline{\omega^k}}_i =
        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \omega^k \, d\omega \right]
        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \, d\omega \right]^{-1},
\end{equation}
%
the Fisher matrix for the signal in the $i$th detector is
%
\begin{equation}
    \mathcal{I}_i = \left(
        \begin{array}{cc}
            \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i} &
            \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_\mathrm{in}} \\
            {\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_\mathrm{in}}}^\intercal &
            {\rho_i}^2 \mathcal{I}_{\boldsymbol\theta_\mathrm{in},\boldsymbol\theta_\mathrm{in}}
        \end{array}
    \right)
\end{equation}
%
where
%
\begin{equation}\label{eq:fisher-matrix}
    \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i} = \bordermatrix{
        ~ & \rho_i & \gamma_i & \tau_i \cr
        \rho_i & 1 & 0 & 0 \cr
        \gamma_i & 0 & {\rho_i}^2 & -{\rho_i}^2 {\overline{\omega}}_i \cr
        \tau_i & 0 & -{\rho_i}^2 {\overline{\omega}}_i & {\rho_i}^2 {\overline{\omega^2}}_i
    }.
\end{equation}
%
(See also \cite{Grover:2013}.) To lowest post\nobreakdashes-Newtonian order, the intrinsic parameters affect the phase evolution but not the amplitude evolution, so for BNS systems we can assume $\mathcal{I}_{\rho_i,\boldsymbol\theta_\mathrm{in}} \approx 0$.
%
\marginpar{Show this?}%
%
When we form the Fisher matrix for the whole network, if the detectors all have the same noise \acp{PSD}, $S_1(\omega) = S_2(\omega) = \cdots = S_n(\omega) \equiv S(\omega)$, then we can change variables from phases and times on arrival $\left\{\gamma_i, \tau_i\right\}_i$ to appropriate differences between them, in order to cancel out all off-diagonal terms between the intrinsic and extrinsic parameters. Because all existing ground-based GW detectors have similar bandpasses, we restrict all of our further attention to just the extrinsic part of the Fisher matrix, $\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i}$.
